---
title: "Patton & Guimond Final Project: EVAULATING TEMPERATURE TRENDS AT US ARMY BASIC TRAINING INSTALLATIONS"
author: "Erik Patton & Austin Guimond"
date: "Fall 2022"
output:
  html_document: default
  pdf_document: default
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory

# Load your packages
#install.packages('TinyText')
#install.packages('LaTex')
#library(LaTex)
#install.packages('agricolae')
#install.packages('corrplot')
#install.packages('sf')
#install.packages('mapview')
#install.packages("viridis")
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(trend)
library(Kendall)
library(tseries)
library(dplyr)
library(formatR)
library(agricolae)
library(corrplot)
library(mapview)
library(sf)
library(leaflet)
library(viridis)
library(cowplot)
library(xtable)
# Set your ggplot theme


# Load temperature data sets
Jackson_Raw <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Raw Data/Jackson_TenYear.csv",
                         stringsAsFactors = TRUE)
KFSI_Clean <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Processed Data/KFSI_Clean.csv",
                         stringsAsFactors = TRUE)
KLSF_Clean <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Processed Data/KLSF_Clean.csv",
                         stringsAsFactors = TRUE) 
KTBN_Clean <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Processed Data/KTBN_Clean.csv",
                         stringsAsFactors = TRUE) 

#Load data sets for mapping visualization
All_Installations_sf <- st_read('/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/Duke/Data_Dissertation/Map Data/tl_2019_us_mil/tl_2019_us_mil.shp')%>%
  select('FULLNAME','geometry')%>%
  filter(FULLNAME=="Ft Jackson"|FULLNAME=="Ft Benning"|FULLNAME=="Ft Sill"|FULLNAME=="Ft Leonard Wood")
All_Counties <- st_read('/Users/erikpatton/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Raw Data/cb_2018_us_county_20m')

```


# Rationale and Research Questions

## Rationale
This project advances research done as part of a dissertation by analyzing temperature trends against thresholds set to protect against heat illness during physical activity. Although it focuses on temperature on military installations and uses temperature thresholds derived by the military, there is broad applicability to any group that performs physically demanding outdoor work. 

##Research Questions
###Research Question #1: Has there been a change to temperature trends at the four army basic training installations in the decade between 2012-2022? 

-1a: Are trends the same or different for Heat Index and Wet Bulb Globe Temperature? 

-1b: Are there trends in the number of “Black Flag” and “Red Flag” days at the installations? 

###Research Question #2: How many days was the temperature in each of the “Color Flag” temperature categories for each installation? 

\newpage

# Dataset Information

Four datasets were obtained from the US Air Force 14th Weather Squadron, which collects “authoritative climate data” for the military. Each dataset contains observations from a weather gauge on or near one of four army basic training installations (Table 1).  

 |      | Fort Jackson, SC    |Fort Benning, GA   |Fort Sill, OK    |Fort Leonard Wood, MO|
 |:---    |   :---:   |   :---:   |     :---:           |   :---:       |   :---:           |
 |Installation Abbreviation   |FJSC|FBGA|FSOK|FLW|
 |Weather Station ID|KCUB|KLSF|KFSI|KTBN|
 |OBSERVATIONS|93,099|93,016|91,404|92,202|

Hourly data for the period of 2012-01-01 00:00:00 to 2022-09-30 23:00 was provided for each basic training installation. Data included the following observations: Temperature (F), Dewpoint (F), Relative Humidity, Heat Index (F), Derived Wet Bulb Glode Temperature (F). Each of the files contained over 91,000 observations, with the individual file’s number of observations dependent on the number of missing observations.  

```{r HOURLY DATA HEADER}
#example of the first ten weather observations made at Ft Jackson, SC.
head(Jackson_Raw, n=10)
```
Publicly releasable version of shapefile data for military installations was obtained from the Defense Installations Spatial Data Infrastructure portal. This file includes data on nine attributes (variables) for 859 military installations (observations) within United States territory. The data file was wrangled using the dyplr pipe function to first select for two columns, ‘FULLNAME’ and ‘geometry’, then filtering by ‘FULLNAME’ for each of the four installations of interest. This resulted in four 1x2 size data frames, one for each installation, containing installation name and multipolygon geometry.  

```{r}
#inspect the features in the Installation shape file data.
(All_Installations_sf)
```
2018 county data at 20-meter resolution was obtained from the US Census Bureau and used to build State boundary maps. This file contains data for ten attributes (variables) for 3,220 counties (observations) within the United States. A data frame for each State with a basic training installation was created by filtering this dataset using the appropriate STATEFP, resulting in a data frame for each State of 10 variables and the number of observations corresponding to the number of counties in the State. 

```{r EXAMPLE SHOWING SOUTH CAROLINA}
#filtering data and creating State maps; showing SC as the example.
GA_counties_sf<- All_Counties %>% 
  filter(STATEFP == 13) 

MO_counties_sf<- All_Counties %>% 
  filter(STATEFP == 29) 

SC_counties_sf<- All_Counties %>% 
  filter(STATEFP == 45) 

OK_counties_sf<- All_Counties %>% 
  filter(STATEFP == 40) 

head(SC_counties_sf)
mapview(SC_counties_sf)
```


\newpage

# Exploratory Analysis 
## Exploratory Analysis Question #1a: Are trends the same or different for Heat Index and Wet Bulb Globe Temperature? 
```{r TEMPORARY ONLY, OLD CODE, USED TO ASSIGN VARIABLES NEEDED FOR QUESTION 1B}

#Load Jackson Data
Jackson_Raw <- read.csv("~/Library/Mobile Documents/com~apple~CloudDocs/R Files/Patton_Guimond_ENV872_Final_Project/Data/Raw Data/Jackson_TenYear.csv",
                         stringsAsFactors = TRUE)
Jackson_Wrangle <- Jackson_Raw %>%
  select(YEAR, MO, DAY, WBGTC) %>%
  mutate('date' = make_date(year = YEAR, month = MO, day = DAY))

#Set as date
Jackson_Wrangle$date <- as.Date(Jackson_Wrangle$date, format = "%y/%m/%d")

#Group by date and find max daily temperature
FJSC_Daily_High <- Jackson_Wrangle %>%
  group_by(date) %>% 
  dplyr::summarize(value = max (WBGTC)) %>% 
  as.data.frame()
#Plot max temperatures over time
MaxWBTG_Plot <- ggplot(FJSC_Daily_High, aes(x = date, y = value)) +
  geom_line()+
  geom_smooth(method=lm, col= 'red')+
 ggtitle("Max Wet Bulb Over Time")+
  xlab("Time (Days)") + ylab("Wet Bulb Globe Temperature")
print(MaxWBTG_Plot)

#Look for NA values in data and remove
summary(FJSC_Daily_High$value)
FJSC_Clean_WetBulb <- 
  FJSC_Daily_High %>% 
  mutate(FJSC_WBGTC_Clean = zoo::na.approx(value))
summary(FJSC_Clean_WetBulb)

#Filter for date and NA omitted Temp
FJSC_Max_WetBulb <- FJSC_Clean_WetBulb%>%
  select(date, FJSC_WBGTC_Clean)
summary(FJSC_Max_WetBulb)

#Create Time series object and decompose

FJSC_Daily_High_ts <- ts(FJSC_Max_WetBulb$FJSC_WBGTC_Clean, start = c(2012,01,01), frequency = 365)
FJSC_Daily_High_decomp <- stl(FJSC_Daily_High_ts,s.window = "periodic")
plot(FJSC_Daily_High_decomp)

FJSC_Daily_WetBulb_Trend <- Kendall::SeasonalMannKendall(FJSC_Daily_High_ts)
summary(FJSC_Daily_WetBulb_Trend)

#Subtract seasonality and run seasonally adjusted Mann Kendall
FJSC_Daily_Components <- as.data.frame(FJSC_Daily_High_decomp$time.series[,1:3])

FJSC_Daily_Components <- mutate(FJSC_Daily_Components,
        WBGTC = FJSC_Max_WetBulb$FJSC_WBGTC_Clean,     
        Date = FJSC_Max_WetBulb$date)

FJSC_WetBulb_SeasonAdj <- FJSC_Daily_Components %>%
  mutate(Subtract.Season = FJSC_Daily_Components$WBGTC - FJSC_Daily_Components$seasonal)
summary(FJSC_WetBulb_SeasonAdj)

FJSC_NonSeasonal_WetBulb_Trend <- Kendall::MannKendall(FJSC_WetBulb_SeasonAdj$Subtract.Season)
summary(FJSC_NonSeasonal_WetBulb_Trend)
```

## Exploratory Analysis Question #1b: Are there trends in the number of “Black Flag” and “Red Flag” days at the installations? 
"Black Flag" refers to days when the local WBGT is >90 Fahrenheit. At this temperature, training on military installations is significantly curtailed. "Red Flag" days refer to WBGT  between 88 and 90 degrees Fahrenheit. To answer a question about numbers and trends for "Black" and "Red" flag days, we first created a table with the standard for each color level and called this table "Hazardous_Days_Limits". Within this table we also created columns to translate degrees Fahrenheit to degrees Celsius to better standardize the analysis.
```{r,include=FALSE}
#Load the temperature thresholds
HD_First_Column <- c("None","Green","Yellow","Red","Black")
HD_Temp_Min <- c(32,82,85,88,90)
HD_Temp_Min <- as.numeric(HD_Temp_Min)
HD_Temp_Min <- signif(HD_Temp_Min,digits=2)
HD_Temp_Max <- c(81.999,84.999,87.999,89.999,104)
HD_Temp_Max <- as.numeric(HD_Temp_Max)
HD_Temp_Max <- signif(HD_Temp_Max,digits=2)
Hazardous_Days_Limits <- cbind(HD_First_Column,HD_Temp_Min,HD_Temp_Max)
Hazardous_Days_Limits <- as.data.frame(Hazardous_Days_Limits)
Hazardous_Days_Limits$HD_Temp_Min <- as.integer(Hazardous_Days_Limits$HD_Temp_Min)
Hazardous_Days_Limits$HD_Temp_Max <- as.integer(Hazardous_Days_Limits$HD_Temp_Max)
```
```{r}
Hazardous_Days_Limits <-  Hazardous_Days_Limits%>%
  mutate(HD_Temp_Min_C = (HD_Temp_Min-32)*(5/9))%>%
  mutate(HD_Temp_Max_C = (HD_Temp_Max-32)*(5/9))%>%
  rename(Flag_Color=HD_First_Column, Temp_Limit_Max_F=HD_Temp_Max,Temp_Limit_Min_F=HD_Temp_Min, Temp_Limit_Min_C=HD_Temp_Min_C, Temp_Limit_Max_C=HD_Temp_Max_C)
Hazardous_Days_Limits
```
Next, filtering using data from question #1a was run to display just Black flag days and Red flag + Black flag days. Red flag days were not considered independent of Black flag days because an analysis of high temperature weather trends should not look at only an intermediate temperature band, since any day that exceeded WBGT >90F also was a Red flag day, having first passed through this temperature range (88-90F). Furthermore, any training restrictions or modification put in place during a Red flag day are also in place on Black flag days. 

The Fort Jackson, SC results are displayed as a representative sample of the output from the four installations. 
```{r FORT JACKSON, SC FLAG COUNT}
#Count number of days in each flag threshold. NOTE VARIABLES WILL HAVE TO BE VERIFIED WITH WHAT IS USED IN QUESTION 1A, POSSIBILY USING OLDER VERSION
FJSC_Total.BlackFlag <- sum(FJSC_Max_WetBulb$FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FJSC_Total.RedFlag <- sum(FJSC_Max_WetBulb$FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[4] & FJSC_Max_WetBulb$FJSC_WBGTC_Clean < Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FJSC_Total.YellowFlag <- sum(FJSC_Max_WetBulb$FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[3] & FJSC_Max_WetBulb$FJSC_WBGTC_Clean < Hazardous_Days_Limits$Temp_Limit_Min_C[4])
FJSC_Total.GreenFlag <- sum(FJSC_Max_WetBulb$FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[2] & FJSC_Max_WetBulb$FJSC_WBGTC_Clean < Hazardous_Days_Limits$Temp_Limit_Min_C[3])
FJSC_Total.NoFlag <- sum(FJSC_Max_WetBulb$FJSC_WBGTC_Clean < Hazardous_Days_Limits$Temp_Limit_Min_C[2])
FJSC_Total.AllFlags <- rbind(FJSC_Total.BlackFlag,FJSC_Total.RedFlag,FJSC_Total.YellowFlag,FJSC_Total.GreenFlag,FJSC_Total.NoFlag)
FJSC_Total.AllFlags <- as.data.frame(FJSC_Total.AllFlags)
FJSC_Total.AllFlags <- cbind(Hazardous_Days_Limits[1],FJSC_Total.AllFlags)
FJSC_Total.AllFlags$Flag_Color <- rev(FJSC_Total.AllFlags$Flag_Color) 
#reverse the color order to match the correct number of days
FJSC_Total.AllFlags <- rename(FJSC_Total.AllFlags, Number_Days=V1)
FJSC_Total.AllFlags <- FJSC_Total.AllFlags%>%
    mutate(Installation = "FJSC")%>%
  select(Installation,Flag_Color,Number_Days)
#view(FJSC_Total.AllFlags)
```

```{r, include=FALSE} 
#TEMP WORKSPACE DETELE IN FINAL --- ONLY DONE TO CREATE CLEAN VARIABLES FOR MAKING CELCIUS AND COUNTING FLAG DAYS IN SUBSEQUENT CODE FOR QUESTIONS 1B AND 2
#created to wrangle KFSI data into degrees C and remove all NAs. Will be replaced by code that Austin inputs for question 1A.
KFSI_Wrangle <- KFSI_Clean %>%
  select(Year, Month, Day, Derived.Wet.Bulb.Globe.Temperature..F.) %>%
  mutate('date' = make_date(year = Year, month = Month, day = Day))

#Set as date
KFSI_Wrangle$date <- as.Date(KFSI_Wrangle$date, format = "%y/%m/%d")
KFSI_Wrangle_Update <- KFSI_Wrangle %>%
  select(date, Derived.Wet.Bulb.Globe.Temperature..F.)

#Group by date and find max daily temperature
Daily_High_KFSI <- KFSI_Wrangle_Update %>%
  group_by(date) %>% 
  dplyr::summarize(value = max (Derived.Wet.Bulb.Globe.Temperature..F.)) %>% 
  as.data.frame()

#Look for NA values in data and remove. Create date range for avalaible data
summary(Daily_High_KFSI$value)
KFSI_Filtered_Date <- Daily_High_KFSI %>%
  filter(between(date, as.Date("2012-01-10"), as.Date("2022-09-30")))
#estimate NA values by average
Clean_Max_WetBulb_KSFI <- 
  KFSI_Filtered_Date %>% 
  mutate(Wet_Bulb = zoo::na.approx(value))

Clean_Max_WetBulb_KSFI <- Clean_Max_WetBulb_KSFI %>%
  select(date,Wet_Bulb)%>%
  drop_na()
summary(Clean_Max_WetBulb_KSFI)


#FT BENNING
KLSF_Wrangle <- KLSF_Clean %>%
  select(Year, Month, Day, Derived.Wet.Bulb.Globe.Temperature..F.) %>%
  mutate('date' = make_date(year = Year, month = Month, day = Day))

#Set as date
KLSF_Wrangle$date <- as.Date(KLSF_Wrangle$date, format = "%y/%m/%d")
KLSF_Wrangle_Update <- KLSF_Wrangle %>%
  select(date, Derived.Wet.Bulb.Globe.Temperature..F.)

#Group by date and find max daily temperature
Daily_High_KLSF <- KLSF_Wrangle_Update %>%
  group_by(date) %>% 
  dplyr::summarize(value = max (Derived.Wet.Bulb.Globe.Temperature..F.)) %>% 
  as.data.frame()

#Look for NA values in data and remove. Create date range for avalaible data
summary(Daily_High_KLSF$value)
KLSF_Filtered_Date <- Daily_High_KLSF %>%
  filter(between(date, as.Date("2012-01-10"), as.Date("2022-09-30")))
#estimate NA values by average
Clean_Max_WetBulb_KLSF <- 
  KLSF_Filtered_Date %>% 
  mutate(Wet_Bulb = zoo::na.approx(value))

Clean_Max_WetBulb_KLSF <- Clean_Max_WetBulb_KLSF %>%
  select(date,Wet_Bulb)%>%
  drop_na()
summary(Clean_Max_WetBulb_KLSF)


#FT LEONARD WOOD
KTBN_Wrangle <- KTBN_Clean %>%
  select(Year, Month, Day, Derived.Wet.Bulb.Globe.Temperature..F.) %>%
  mutate('date' = make_date(year = Year, month = Month, day = Day))

#Set as date
KTBN_Wrangle$date <- as.Date(KTBN_Wrangle$date, format = "%y/%m/%d")
KTBN_Wrangle_Update <- KTBN_Wrangle %>%
  select(date, Derived.Wet.Bulb.Globe.Temperature..F.)

#Group by date and find max daily temperature
Daily_High_KTBN <- KTBN_Wrangle_Update %>%
  group_by(date) %>% 
  dplyr::summarize(value = max (Derived.Wet.Bulb.Globe.Temperature..F.)) %>% 
  as.data.frame()

#Look for NA values in data and remove. Create date range for avalaible data
summary(Daily_High_KTBN$value)
KTBN_Filtered_Date <- Daily_High_KTBN %>%
  filter(between(date, as.Date("2012-01-10"), as.Date("2022-09-30")))
#estimate NA values by average
Clean_Max_WetBulb_KTBN <- 
  KTBN_Filtered_Date %>% 
  mutate(Wet_Bulb = zoo::na.approx(value))

Clean_Max_WetBulb_KTBN <- Clean_Max_WetBulb_KTBN %>%
  select(date,Wet_Bulb)%>%
  drop_na()
#summary(Clean_Max_WetBulb_KTBN)

```

```{r ALL OTHER INSTALLATIONS FLAG DAY COUNT}
#FORT SILL, OK
#Rework this so that FSOK_Clean_C is referencing the correct data set
FSOK_Clean_C <- Clean_Max_WetBulb_KSFI%>%
  mutate(Wet_Bulb_C = (Wet_Bulb-32)*(5/9))

FSOK_Total.BlackFlag <- sum(FSOK_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FSOK_Total.RedFlag <- sum(FSOK_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4] & FSOK_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FSOK_Total.YellowFlag <- sum(FSOK_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[3] & FSOK_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[4])
FSOK_Total.GreenFlag <- sum(FSOK_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[2] & FSOK_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[3])
FSOK_Total.NoFlag <- sum(FSOK_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[2])
FSOK_Total.AllFlags <- rbind(FSOK_Total.BlackFlag,FSOK_Total.RedFlag,FSOK_Total.YellowFlag,FSOK_Total.GreenFlag,FSOK_Total.NoFlag)
FSOK_Total.AllFlags <- as.data.frame(FSOK_Total.AllFlags)%>%
  cbind(Hazardous_Days_Limits[1],FSOK_Total.AllFlags)
FSOK_Total.AllFlags$Flag_Color <-  rev(FSOK_Total.AllFlags$Flag_Color) 
#reverse the color order to match the correct number of days
FSOK_Total.AllFlags <- rename(FSOK_Total.AllFlags, Number_Days=V1)
FSOK_Total.AllFlags <-  FSOK_Total.AllFlags%>%
    mutate(Installation = "FSOK")%>%
  select(Installation,Flag_Color,Number_Days)
#view(FSOK_Total.AllFlags)


#FORT BENNING, GA
#CLEAN_C WILL NEED TO BE REWORKED ONCE FINAL VARIABLES ARE ASSIGNED
FBGA_Clean_C <- Clean_Max_WetBulb_KLSF%>%
  mutate(Wet_Bulb_C = (Wet_Bulb-32)*(5/9))

FBGA_Total.BlackFlag <- sum(FBGA_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FBGA_Total.RedFlag <- sum(FBGA_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4] & FBGA_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FBGA_Total.YellowFlag <- sum(FBGA_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[3] & FBGA_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[4])
FBGA_Total.GreenFlag <- sum(FBGA_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[2] & FBGA_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[3])
FBGA_Total.NoFlag <- sum(FBGA_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[2])
FBGA_Total.AllFlags <- rbind(FBGA_Total.BlackFlag,FBGA_Total.RedFlag,FBGA_Total.YellowFlag,FBGA_Total.GreenFlag,FBGA_Total.NoFlag)
FBGA_Total.AllFlags <- as.data.frame(FBGA_Total.AllFlags)%>%
  cbind(Hazardous_Days_Limits[1],FBGA_Total.AllFlags)
FBGA_Total.AllFlags$Flag_Color <-  rev(FBGA_Total.AllFlags$Flag_Color) 
#reverse the color order to match the correct number of days
FBGA_Total.AllFlags <- rename(FBGA_Total.AllFlags, Number_Days=V1)
FBGA_Total.AllFlags <-  FBGA_Total.AllFlags%>%
    mutate(Installation = "FBGA")%>%
  select(Installation,Flag_Color,Number_Days)
#view(FBGA_Total.AllFlags)

#FT LEONARD WOOD, MO
#CLEAN_C WILL NEED TO BE REWORKED ONCE FINAL VARIABLES ARE ASSIGNED
FLW_Clean_C <- Clean_Max_WetBulb_KTBN%>%
  mutate(Wet_Bulb_C = (Wet_Bulb-32)*(5/9))
FLW_Total.BlackFlag <- sum(FLW_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FLW_Total.RedFlag <- sum(FLW_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4] & FLW_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[5])
FLW_Total.YellowFlag <- sum(FLW_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[3] & FLW_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[4])
FLW_Total.GreenFlag <- sum(FLW_Clean_C$Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[2] & FLW_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[3])
FLW_Total.NoFlag <- sum(FLW_Clean_C$Wet_Bulb_C < Hazardous_Days_Limits$Temp_Limit_Min_C[2])
FLW_Total.AllFlags <- rbind(FLW_Total.BlackFlag,FLW_Total.RedFlag,FLW_Total.YellowFlag,FLW_Total.GreenFlag,FLW_Total.NoFlag)
FLW_Total.AllFlags <- as.data.frame(FLW_Total.AllFlags)%>%
  cbind(Hazardous_Days_Limits[1],FLW_Total.AllFlags)
FLW_Total.AllFlags$Flag_Color <-  rev(FLW_Total.AllFlags$Flag_Color) 
#reverse the color order to match the correct number of days
FLW_Total.AllFlags <- rename(FLW_Total.AllFlags, Number_Days=V1)
FLW_Total.AllFlags <-  FLW_Total.AllFlags%>%
  mutate(Installation = "FLW")%>%
  select(Installation,Flag_Color,Number_Days)
#view(FLW_Total.AllFlags)
```

```{r CREATE FINAL TABLE WITH ALL FLAG DAYS}
All_Installations.AllFlags <- rbind(FJSC_Total.AllFlags,FBGA_Total.AllFlags,FSOK_Total.AllFlags,FLW_Total.AllFlags)
view(All_Installations.AllFlags)
```

### Trend Alanysis for Black and Red Flag Days
```{r, include=FALSE}
#THIS CODE CHUNK SHOULD NOT BE INCLUDED IN THE FINAL PROJECT
#creating subsets for each installation with only the daily high temperature at or exceeding Red Flag threshold - by summer months
FJSC_B_R_subset <- FJSC_Max_WetBulb%>%
  subset(FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year,month)%>%
  count(month)%>%
  rename(FJSC_Count_Days_Red_and_Above=n)
view(FJSC_B_R_subset)


FSOK_B_R_subset <- FSOK_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year,month)%>%
  count(month)%>%
  rename(FSOK_Count_Days_Red_and_Above=n)

FBGA_B_R_subset <- FBGA_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year,month)%>%
  count(month)%>%
  rename(FBGA_Count_Days_Red_and_Above=n)

FLW_B_R_subset <- FLW_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year,month)%>%
  count(month)%>%
  rename(FLW_Count_Days_Red_and_Above=n)

#creating combined "Red and Black" subset
R_B_subset <- full_join(FSOK_B_R_subset,FJSC_B_R_subset)
R_B_subset <- full_join(FBGA_B_R_subset,R_B_subset)
R_B_subset <- full_join(FLW_B_R_subset,R_B_subset)
R_B_subset <- R_B_subset%>%
  mutate('date' = make_date(year = year, month = month))%>%
  filter(month == 6||month==7||month==8)%>%#filters for June, July, August
view(R_B_subset)
```

```{r}
#Ft Jackson wrangling for Red and Black flag trends, shown as example of how all installations were done
FJSC_B_R_subset2 <- FJSC_Max_WetBulb%>%
  subset(FJSC_WBGTC_Clean >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year)%>%
  count(year)%>%
  rename(FJSC_Count_Days_Red_and_Above=n)
view(FJSC_B_R_subset2)
```

```{r, include=FALSE}
#other three installations, not to display in final document
FSOK_B_R_subset2 <- FSOK_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year)%>%
  count(year)%>%
  rename(FSOK_Count_Days_Red_and_Above=n)

FBGA_B_R_subset2 <- FBGA_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year)%>%
  count(year)%>%
  rename(FBGA_Count_Days_Red_and_Above=n)

FLW_B_R_subset2 <- FLW_Clean_C%>%
  subset(Wet_Bulb_C >= Hazardous_Days_Limits$Temp_Limit_Min_C[4])%>%
  mutate(month = month(date))%>%
  mutate(year=year(date))%>%
  group_by(year)%>%
  count(year)%>%
  rename(FLW_Count_Days_Red_and_Above=n)
view(FLW_B_R_subset2)

#creating combined "Red and Black" subset
R_B_subset2 <- full_join(FSOK_B_R_subset2,FJSC_B_R_subset2)
R_B_subset2 <- full_join(FBGA_B_R_subset2,R_B_subset2)
R_B_subset2 <- full_join(FLW_B_R_subset2,R_B_subset2)
#R_B_subset2 <- R_B_subset2
  #mutate('date' = make_date(year = year, month = month))%>%
  #filter(month == 6||month==7||month==8)%>%#filters for June, July, August
view(R_B_subset2)
```

```{r, include=FALSE}
R_B_subset_plot2 <- ggplot(R_B_subset2,aes(x=year))+
    ylab("Days At RED or Above WBGT Conditions")+
  scale_x_continuous(name="Year",n.breaks=11)+
  theme(axis.text.x = element_text(size=7, angle=45), plot.title=element_text(hjust=0.5, vjust=0.5, face='bold'))+ 
  geom_point(aes(y=FLW_Count_Days_Red_and_Above),color='black',shape=0)+
  geom_line(aes(y=FLW_Count_Days_Red_and_Above),color='red')+
  geom_smooth(aes(y=FLW_Count_Days_Red_and_Above),se=FALSE,color='red')+
  geom_point(aes(y=FBGA_Count_Days_Red_and_Above),color='black',shape=1)+
  geom_line(aes(y=FBGA_Count_Days_Red_and_Above),color='blue')+
  geom_smooth(aes(y=FBGA_Count_Days_Red_and_Above), se=FALSE,color='blue')+
  geom_point(aes(y=FSOK_Count_Days_Red_and_Above),color='black',shape=2)+
  geom_line(aes(y=FSOK_Count_Days_Red_and_Above),color='yellow')+
  geom_smooth(aes(y=FSOK_Count_Days_Red_and_Above), se=FALSE,color='yellow')+
  geom_point(aes(y=FJSC_Count_Days_Red_and_Above),color='black',shape=4)+
  geom_line(aes(y=FJSC_Count_Days_Red_and_Above),color='white')+
  geom_smooth(aes(y=FJSC_Count_Days_Red_and_Above), se=FALSE,color='white')
 
R_B_subset_plot2
```
To explore the trend analysis for the number of Black flag days or Red flag + Black flag days, data was wrangled by grouping by (year, month) and then counting the number of days with a WBGT value greater than or equal to the lower limit for Red flag day (>= 88F). The results were first put into the following graph to visual inspection:
```{r, collpase=TRUE}
R_B_subset_plot2
```
There is no obvious trend to the number of WBGT days, either increasing or decreasing, among the installations across the decade studied. To determine statistically if a correlation exists, "corrtest" was conducted between the year and the number of WBGT days.
```{r, include=FALSE}
FJSC_CorTest <- cor.test(R_B_subset2$FJSC_Count_Days_Red_and_Above,R_B_subset2$year)
FJSC_CorTest
FBGA_CorTest <- cor.test(R_B_subset2$FBGA_Count_Days_Red_and_Above,R_B_subset2$year)
FBGA_CorTest
FSOK_CorTest <- cor.test(R_B_subset2$FSOK_Count_Days_Red_and_Above,R_B_subset2$year)
FSOK_CorTest
FLW_CorTest <- cor.test(R_B_subset2$FLW_Count_Days_Red_and_Above,R_B_subset2$year)
FLW_CorTest
```
Correlation testing, using the Pearson's product-moment correlation, found no statistically significant p-values when measuring the difference in the number of days at Red flag or above and the year. Three of the installations (Ft Benning, Ft Jackson, and Ft Sill) did have positive correlations, although with confidence intervals that extended from negative values to positive values. The fourth installation, Ft Leonard Wood, actually had a slight negative correlation, implying a reduction in the number of days at or above Red flag, although this correlation was also not statistically significant. Interestingly, this negative correlation had the strongest statistical significance, with a p-value of 0.2, although this is still far from the 0.05 p-value that would allow a rejection of the null hypothesis. The results for Ft Jackson, SC and Ft Leonard Wood, MO are displayed below:
```{r, collapse=TRUE}
FJSC_CorTest
FLW_CorTest
```


## Exploratory Analysis Question #2: How many days was the temperature in each of the “Color Flag” temperature categories for each installation?
To determine the number of days that each installation was in a heat category, code from the Black and Red Flag trend analysis was used. After wrangling to ensure the WBGT temperature was in degrees C and removing any rows with NAs in WBGT column, each row was compared to the "Hazardous_Days_Limits" table that was created at the start of this project. 

The total number of Black flag days at Ft Jackson over the time period studied was:
```{r, collapse=TRUE}
FJSC_Total.BlackFlag
```
and the total number of Black plus Red flag days was:
```{r, collapse=TRUE}
FJSC_Total.BlackFlag+FJSC_Total.RedFlag
```

The final output is the following table with Black and Red flag day numbers for each installation:
```{r, Include=FALSE}
AllFlags_Table <- subset(All_Installations.AllFlags, Flag_Color %in% c("Black","Red"))
AllFlags_Table
```

\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 
In terms of real number of days exceeding the Black or Red flag thresholds it is clear that Ft Benning, GA experienced the most high WBGT days. In total, there were 772 days at FBGA that were > 88F. Ft Sill, OK, and Ft Jackson, SC tied with a combined total of 510 Black or Red flag days, although Ft Sill, OK experienced slightly more of the hottest days (232 Black flags days at Ft Sill vs 217 Black flag days at Ft Jackson). Ft Leonard Wood, MO had the least amount of high heat training days, at 230. 

Ft Benning experiencing the greatest number of high heat days is expected, as it is the furthest south of all training installations considered in this study, not near a coast, and is located in the humid Southeast. The 4th National Assessment on climate change states that, although the Southeast is warming relatively slowly compared to much of the world, three of the top five large cities experiencing increases in heat wave duration and intensity are located in the Southeast. The number of nights above 75F in the Southeast has increased at roughly double the historical average, an important consideration during days-long field training exercises during which a soldier's body relies on cooler nighttime temperatures to "reset" and prepare for another hot day. ADD IN SOMETHING HERE ABOUT TRENDS ABOVE BLACK AND RED?

```{r}
#FSOK = Fort Sill, Oklahoma
FSOK_KFSI_sfg <- st_point(c(-98.391,34.637))
FSOK_KFSI_sfc <- st_sfc(FSOK_KFSI_sfg, crs=4326)
FSOK_KFSI_sf <- st_as_sf(FSOK_KFSI_sfc)
FSOK_KFSI_sf$Name = 'Ft Sill, OK'

#FJSC = Fort Jackson, SC
FJSC_KCUB_sfg <- st_point(c(-81.01,33.99))
FJSC_KCUB_sfc <- st_sfc(FJSC_KCUB_sfg,crs=4326)
FJSC_KCUB_sf <- st_as_sf(FJSC_KCUB_sfc)
FJSC_KCUB_sf$Name = 'Ft Jackson, SC'

#FLW = Fort Leaonard Wood, MO
FLW_KTBN_sfg <- st_point(c(-92.12,37.75))
FLW_KTBN_sfc <- st_sfc(FLW_KTBN_sfg,crs=4326)
FLW_KTBN_sf <- st_as_sf(FLW_KTBN_sfc)
FLW_KTBN_sf$Name <- 'Ft Leonard Wood, MO'

#FBGA = Fort Benning, GA
FBGA_KLSF_sfg <- st_point(c(-84.97,32.35))
FBGA_KLSF_sfc <- st_sfc(FBGA_KLSF_sfg,crs=4326)
FBGA_KLSF_sf <- st_as_sf(FBGA_KLSF_sfc)
FBGA_KLSF_sf$Name <- 'Ft Benning, BA'

#All on the same MapView
Guage_Locations_df <- rbind(FBGA_KLSF_sf,FLW_KTBN_sf,FJSC_KCUB_sf,FSOK_KFSI_sf)
#basic_installation_map <- mapview(Guage_Locations_df)
#print(basic_installation_map)

#Individual Installations
FBGA_Installations_sf <- All_Installations_sf%>%
  select('FULLNAME','geometry')%>%
  filter(FULLNAME=="Ft Benning")
FJSC_Installations_sf <- All_Installations_sf%>%
  select('FULLNAME','geometry')%>%
  filter(FULLNAME=="Ft Jackson")
FSOK_Installations_sf <- All_Installations_sf%>%
  select('FULLNAME','geometry')%>%
  filter(FULLNAME=="Ft Sill")
FLW_Installations_sf <- All_Installations_sf%>%
  select('FULLNAME','geometry')%>%
  filter(FULLNAME=="Ft Leonard Wood")

mapview(All_Installations_sf, zcol="FULLNAME",layer_name="Basic Training Installations",alpha=1,legend=FALSE)+
  mapview(Guage_Locations_df, cex=3, layer.name="Installation Gauge Locations")
```


Ft Sill, OK and Ft Jackson, SC, sharing the #2 rank was surprising. 



\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
